---
title: "Public Housing Outreach"
author: "Christopher Prener, Ph.D. and Branson Fox"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
# load dependencies
library(dplyr)
library(here)
library(leaflet)
library(RColorBrewer)
library(readr)
library(sf)
library(tigris)

# load data
housing <- st_read(here("data", "clean2", "public_housing.geojson"), crs = 4326,
                       stringsAsFactors = FALSE) %>%
  filter(county %in% c("St. Louis", "City of St. Louis"))
                       
counties <- counties(state = 29, class = "sf") %>%
  select(GEOID) %>%
  filter(GEOID %in% c("29189", "29510")) %>%
  st_transform(crs = 4326)

# create palette
pal <- brewer.pal(n = 1, name = "Set1")

# clean-up
# rm(focal_zips)

# create spreadsheet output
full <- housing
st_geometry(full) <- NULL
write_csv(full, "outreach_public_housing.csv")
```

## Background
This interactive report contains the locations of public housing complexes in St. Louis City and County. These data were obtained from HUD and are provided with no warranty. There may be missing data, and locations may also have been incorrectly geocoded by HUD. Clicking on a complex will reveal details, including the number of units on file with HUD.

## Map
Public housing complexes are labeled with **red circles** (n = `r nrow(housing)`). County boundaries are marked in black.

<br>

```{r map, echo = FALSE, out.width = '100%', out.height='800px'}
# map 
leaflet() %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addPolygons(data = counties,
              color = "#000000",
              opacity = 1,
              fillColor = "transparent") %>%
  addCircleMarkers(data = housing, 
                   radius = 8,
                   opacity = 1,
                   color = pal[1],
                   fillColor = pal[1],
                   popup = paste0("<b>Name</b>: ", housing$title, "<br>",
                                  "<b>Category</b>: Public Housing <br>",
                                  "<b>Authority</b>: ", housing$authority, "<br>",
                                  "<b>Units</b>: ", housing$units, "<br>",
                                  "<b>Address:</b><br>",
                                  housing$address, "<br>",
                                  housing$address2, "<br>",
                                  "<b>County</b>: ", housing$county))
```

<br>

## Acknowledgements
Branson Fox, a Saint Louis University undergraduate student, wrote and executed the initial scripts for scraping bulk location data from HUD.

<br>
